---
title: "DS311 - R Lab Assignment"
author: "Justin Shay"
date: '`r Sys.Date()`'
output:
  html_document:
    theme: united
    highlight: tango
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
nrow(mtcars)
ncol(mtcars)
# Answer:
print("There are total of __11___ variables and ___32__ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are __5___ discrete variables and __6___ continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)
m
v
s
# print(paste("The average of Mile Per Gallon from this data set is ", 20.09 , " with variance ", 36.32 , " and standard deviation", 6.03 , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
average_mpg_cylinder_class <- aggregate(mpg ~ cyl,data = mtcars, FUN = mean )
average_mpg_cylinder_class
standard_deviation_gear_class <- aggregate(mpg ~ gear, data = mtcars, FUN = sd)
standard_deviation_gear_class
```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
crosstab <- table(mtcars$cyl, mtcars$gear)
crosstab
print("The most common car type in this data set is car with __8__ cylinders and __3 __ gears. There are total of __12___ cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
# Load the data set
data("PlantGrowth")

# Boxplot to compare plant weights across groups
boxplot(weight ~ group, 
        data = PlantGrowth,
        main = "Comparison of Plant Weights",
        xlab = "Test Group",
        ylab = "Plant Weight")
    

```

Result:

=> Report a paragraph to summarize your findings from the plot!

The first thing I noticed when observing this box plot was that the trt1 group has plants that are on the smaller size compared to the other groups. From this observation, we can also see that trt2 group has the heavier plants. The crtl group is right in between the two groups. Another observation I had is the trt1 group is more diverse than the others, with the heaviest weighing plant being near 6.0 grams and its lightest weighing plant at almost 3.5 grams. Theres more of an even range in that group. 


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}

hist(mtcars$mpg, breaks = 10, main = "Cars miles per gallon", xlab = "Miles Per Gallon", ylab = "Frequency")

print("Most of the cars in this data set are in the class of ___15___ mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
plot(USArrests$Murder, USArrests$Assault, main = "Murder vs Assault data", xlab = "Murder Rate Per 100,000 group", ylab = "Assault Rate per 100,00 group")

``` 

Result:

=> Report a paragraph to summarize your findings from the plot!

An observation I made is that the dots are very spread out across the graph and not all of them are condensed in one area.By looking at the graph, I see there's a correlation between the murder rate and assault rate. The higher the assault rate, the higher the murder rate. Higher assaults mean higher chances of murders. In the corner near 50 on the y axis there inst alot of murders, but as the assaults go up, the murder also increases. 
  

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='housing.csv', mode='wb')
housingData <- read.csv('housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
summary(housingData$Market.Value.per.SqFt)

hist(housingData$Market.Value.per.SqFt,
     breaks = 20,
     main = "Distribution of Market Value per SqFt", xlab = "Market Value per SqFT",
     ylab = "frequency")

plot(housingData$Year.Built, housingData$Market.Value.per.SqFt,
     main = "Market Value per SqFt compared to Year Built", xlab = "Year Built", ylab = "Market Value per SqFt", col = "blue")
```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
boxplot(Market.Value.per.SqFt ~ Neighborhood, data = housingData, main = "Market value per SqFt in differnet Neighborhoods", xlab = "Neighborhood", ylab = "Market Value per SqFt", col = "green")

ggplot(housingData, aes(x = Boro, y = Market.Value.per.SqFt)) +
  geom_boxplot() +
  labs(title = "Market Value per SqFt by Borough",
       x = "Borough",
       y = "Market Value per SqFt") 

plot(housingData$Year.Built, housingData$Market.Value.per.SqFt,
     main = "Market Value per SqFt compared to Year Built", xlab = "Year Built", ylab = "Market Value per SqFt", col = "blue")

ggplot(housingData, aes(x = Market.Value.per.SqFt)) + geom_histogram(binwidth = 10, fill = "orange")+ labs(title = "Market Value per SqFt Distribution by Borough",
       x = "Market Value per SqFt",
       y = "Count") 


```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!

In this exercise, we explored the data of housing prices in New York and how various factors affect the cities housing market.In the first plot, I wanted to examine the differences in market value per Sq Ft across different neighborhoods.Although the graph isn't that clear, it shows that the market values per Sq ft were relatively the same across neighborhoods. Since it wast the clearest, I decided to compare market value per Sq Ft by boroughs instead, and found out that Manhattan has the highest average market value per Sq ft.Next, I decided to look into the relationship between the year the houses were built and their market value per Sq Ft. From the plot, I noticed a concentration of data points from the 1900s to the 2000s, with few to no dots int the 1800s. This can mean that there was a housing market boom at the start of the 1900s which led to the increase of housing prices. For the final graph I decided to create a cleaner visual for Market value per Sq Ft by Borough. 
